---
title: "lab4"
author: "Natasha Djuric"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install.packages("dslabs")
library(tidyverse)
library(dslabs)
library(ggrepel)

gapminder |> as_tibble() |> 
  head(5)
```

```{r}
#Question 1
gapminder |> 
  arrange(-population,-year) |> 
  filter(country %in% c("China","India","United States","Indonesia","Pakistan")) |> 
  ggplot(aes(x=year,y=population,color=country))+
  geom_point()+
  geom_smooth()
  
```
#Question 2

```{r}
gapminder |> 
  filter(year==2015 & country %in% c("Turkey","Poland","South Korea","Russia","Vietnam","South Africa")) |> 
  arrange(-infant_mortality) |> 
  select(country,infant_mortality) |> 
  knitr::kable()
```

#Question 3

```{r Question 3}
gapminder |> 
  filter(year==2000) |> 
  ggplot(aes(x=gdp/population, y=fertility))+
  geom_point()+
  geom_smooth(method="lm")
```
# Question 4

```{r Question 4}
gapminder |> 
  filter(year==2000) |> 
  ggplot(aes(x=gdp/population, y=fertility))+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~continent,scales="free_y")
```
Europe is the only continent where per capita gdp is positively related with fertility rate

```{r}
eu_2000<-gapminder |> 
  filter(year==2000 & continent=="Europe")

eu_2000|> 
  filter(fertility > 1.5, gdp/population > 20000) |> 
  ggplot(aes(x=gdp/population, y=fertility,color=region))+
  ggrepel::geom_label_repel(aes(label=country))+
  geom_point(data=eu_2000)
```
# Question 5
```{r}
gapminder |> 
  filter(year == 2015) |> 
  group_by(continent) |> 
  summarise(popbillions = sum(population)/10^9)
```

# Question 6
(each country gets its own row, but some countries have larger populations so this needs to be weighted by the population)
```{r}
gapminder |> 
  filter(year == 2015) |> 
  summarise(averagelifeexpectancy = sum(life_expectancy *population)/sum(population)
```

# Question 7
```{r}
gapminder |> 
  filter(year == 2015 & continent %in% c("Europe","Africa")) |> 
  group_by(continent) |> 
  summarise(meanlifeexpectancy = sum(life_expectancy*population)/sum(population)) 
  
library(gapminder)
  
```

# Exercise 2: Use data transformation and visualization to explore the following open-ended question
